
# Finding Datasets of Interest

Goal: RNA-seq, human, recent (last ~3 years), and with sufficient samples.

## RNA-seq series by platform technology

Select distinct sets of series (GSE) and platforms (GPL) where the technology is "high-throughput sequencing", after mapping series to the platforms it involves.

```{r}
sql <- paste(
  "SELECT DISTINCT gse.title, gse.gse, gpl.title",
  "FROM gse JOIN gse_gpl ON gse_gpl.gse = gse.gse",
  "JOIN gpl ON gse_gpl.gpl = gpl.gpl",
  "WHERE gpl.technology = 'high-throughput sequencing'"
)
rs <- dbGetQuery(con, sql)
dim(rs)
```

## Add organism and recency filters

Edited query to also select series for recency and for "Homo sapiens".

```{r}
sql <- paste(
  "SELECT DISTINCT gse.title, gse.gse, gpl.title, gse.submission_date",
  "FROM gse JOIN gse_gpl ON gse_gpl.gse = gse.gse",
  "JOIN gpl ON gse_gpl.gpl = gpl.gpl",
  "WHERE gse.submission_date > '2022-01-01' AND",
  "gpl.organism LIKE '%Homo sapiens%' AND",
  "gpl.technology LIKE '%high-throughput seq%'"
)
rs <- dbGetQuery(con, sql)
dim(rs)
```

## Add a topic keyword (e.g., ovarian)

Edited query to also select for a topic.

```{r}
sql <- paste(
  "SELECT DISTINCT gse.title, gse.gse, gpl.title, gse.submission_date",
  "FROM gse JOIN gse_gpl ON gse_gpl.gse = gse.gse",
  "JOIN gpl ON gse_gpl.gpl = gpl.gpl",
  "WHERE gse.submission_date > '2022-01-01' AND",
  "gse.title LIKE '%ovarian%' AND",
  "gpl.organism LIKE '%Homo sapiens%' AND",
  "gpl.technology LIKE '%high-throughput seq%'"
)
rs <- dbGetQuery(con, sql)
dim(rs)
head(rs[, c('gse','title')])
```

## Heuristic: require a counts-like supplementary file

```{r}
sql <- paste(
  "SELECT DISTINCT gse.title, gse.gse, gpl.title, gse.submission_date, gse.supplementary_file",
  "FROM gse JOIN gse_gpl ON gse_gpl.gse = gse.gse",
  "JOIN gpl ON gse_gpl.gpl = gpl.gpl",
  "WHERE gse.submission_date > '2022-01-01' AND",
  "gse.title LIKE '%ovarian%' AND",
  "gpl.organism LIKE '%Homo sapiens%' AND",
  "gpl.technology LIKE '%high-throughput sequencing%'",
  "ORDER BY gse.submission_date DESC"
)
rs <- dbGetQuery(con, sql)
counts_files <- rs$supplementary_file[grep("count|cnt", rs$supplementary_file, ignore.case = TRUE)]
series_of_interest <- rs$gse[grep("count|cnt", rs$supplementary_file, ignore.case = TRUE)]
# Preview filenames
shortened <- unlist(lapply(counts_files, function(x){
  x <- unlist(strsplit(x, ";"))
  x <- x[grep("count|cnt|txt|csv|xlsx", x, ignore.case = TRUE)]
  tail(unlist(strsplit(x, "/")), n = 1)
}))
head(shortened, 10)
```

## Require a minimum number of samples per series
Count the number of samples in the series, select the ones that have enough.
(THIS CHUNK TAKES REALLY LONG???)

```{r}
num_series <- dbGetQuery(con, paste(
  "select * from gsm where series_id in ('",
  paste(series_of_interest, collapse = "','"), "')", sep = ""))

gse.count <- as.data.frame(table(num_series$series_id))
subset(gse.count, Freq > 6)
```
